---
title: "Using Git and GitHub with R"
author: "Mohammad Rashidul Hashan"
date: "2022-10-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Git

Before installing Git all the packages, R and R studio needs to be updated

Download Git from the Google as per operating system

# Check installation

Put the following on the terminal of the R studio

git --version

The output should show Git version more than 2.3 

## Configure Git

install the following package

```{r}
install.packages("usethis")
library(usethis)

# Using the package configure Git, change the configuration for the git

use_git_config(user.name  = "Mohammad Rashidul Hashan", 
               user.email = "hasandmck66@gmail.com" )

```


## Create a local Git repository

You can create a Git repository in the very beginning whenever a new project

is build upon.

However, there is another way to open if already an existing package 

requires local Git repository

```{r}
library(usethis)
use_git()

```

## Commit the first file

After restoring the git window on the R studio pane, its time to commit the
the files under project. Simply clicking the commit option on the Git pane
will make those files.

If anyone wants to check the committed file they can check out on the 
history. YOu can also add messages for each of the commit you make in the 
commit message window that appears after hitting commit button

## Create GitHub account and token 

To connect Git from R studio with GitHub account, you have to install a
package. You have to create personal access token (PAT) for smooth 
functioning of the affiliated GitHub account. Regular update with the
project and corresponding GitHub account will happen. 

```{r}
library(usethis)

create_github_token()

install.packages("gitcreds")
library(gitcreds)

gitcreds_set()

```

## Connecting Git and GitHub 

Push the local R studio project to the GitHub repository. 

```{r}
library(usethis)

use_github()

```

Pull the GitHub repository to the local R studio project

To do that just open a R studio project by pasting the link from the new
GitHub repository that has been made on GitHub. 

Now to push the committed changes from the local R studio project, just 
click on the push option after committing the changes. It will 
automatically update the GitHub repository with the local changes that 
you have made now.

You have to do it on regular occasion to keep updated on both side.

Whenever any changes are made on the remote repository, pull must be done
before pushing the local repository changes are sent to the remote repo.



